---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data used is from the [Dietary Interview - Total Nutrient Intakes, First Day](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DR1TOT_L.htm).

Each row is responses from a unique candidate

# Data formatting

## Read data

```{r}
library(dplyr)
library(ggplot2)
theme_set(theme_bw())

# Download NHANES August 2021-August 2023 to temporary file    
download.file("https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DR1TOT_L.xpt", tf <- tempfile(), mode="wb")

# Create Data Frame From Temporary File
df <- foreign::read.xport(tf)
```

## Merge data

Next we merge the data with the [demographics dataset](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DEMO_L.htm)

```{r}
# Download NHANES August 2021-August 2023 to temporary file    
download.file('https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DEMO_L.xpt', tf <- tempfile(), mode="wb")

# Create Data Frame From Temporary File
demographics <- foreign::read.xport(tf)
```

## Join data

```{r}
df <- df %>%
  left_join(demographics, by = join_by(SEQN))
```

# Water Intake

First let's look at water intake. The variables are

```
DR1_320Z - Total plain water drank yesterday (gm)
DR1_330Z - Total tap water drank yesterday (gm)
DR1BWATZ - Total bottled water drank yesterday (gm)
```
Total plain water includes both tap water and bottled water.

## Overall Water Intake Per Person

```{r}
# Remove missing
df_water <- df %>% 
  filter(!is.na(DR1BWATZ))

# 
df_water <- df %>%
  
```

## By type

If you read through the documentation, there are two types of water intake: tap water and bottled water.

