---
title: "Untitled"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Navigation

* [Demographics](#demographics)
  * [Average Water Intake Per Person](#average-water-intake-per-person)
  * [Average Water Intake by Gender](#average-water-intake-by-gender)
  * [Water Intake by Age](#water-intake-by-age)
  * [Average Water Intake by Race](#race)
  * [Average Water Intake by Education](#education)
* [Water Intake and BMI](#water-intake-and-bmi)
  * [Linear Relationship](#linear-relationship)
  * [Water Intake and Obsesity](#and-obesity)
  * [Including Confounds](#add-confounds)
* [Mental Health](#mental-health)
  * [Depression](#depression)
  * [Tired or Little Energy](#tired-or-little-energy)

# Data loading and merging

## Read data

```{r}
library(dplyr)
library(ggplot2)
theme_set(theme_bw())

diet <- readRDS('./data/total_dietary_intake.rds')
demographics <- readRDS('./data/demographics.rds')
body_measures <- readRDS('./data/body_measures.rds')
depression <- readRDS('./data/depression.rds')
income <- readRDS('./data/income.rds')
```

## Join data

These are all combined on the total nutrient intake data set so that the analysis centers on respondents' water intake

  * [Dietary Interview - Total Nutrient Intakes, First Day](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DR1TOT_L.htm)
  * [Demographics](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DEMO_L.htm)
  * [Examination data collected for body measures](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/BMX_L.htm) (for BMI)
  * [Questionnaire: Depression](https://wwwn.cdc.gov/Nchs/Data/Nhanes/Public/2021/DataFiles/DPQ_L.htm)
  * Income
  
```{r}
### Merge Data
df <- diet %>%
  left_join(demographics, by = join_by(SEQN)) %>%
  left_join(body_measures, by = join_by(SEQN)) %>%
  left_join(depression, by = join_by(SEQN)) %>%
  left_join(income, by = join_by(SEQN))

### Keep only rows where water intake was measured
df <- df %>%
  filter(!is.na(DR1_320Z))

rm(diet, demographics, body_measures, depression, income)
```

The only variable for water intake we need is `DR1_320Z` - `Total plain water drank yesterday (gm)`, so we can just rename it to "water"

Total plain water includes both tap water and bottled water.

```{r}
df <- df %>%
  rename(water = DR1_320Z)
```

For cleaner analyses and graphing, let's also remove outliers of > 3 standard deviations

```{r}
df <- df %>%
  mutate(water_z = scale(water)) %>%
  filter(water_z < 3)
```

# Demographics

## Average Water Intake Per Person

```{r}
df %>%
  summarize(`Average Water Intake` = mean(water))
```

On the average day, the average American drinks **1087ml** of water.

## Average Water Intake by Gender

```{r}
df$gender <- ifelse(df$RIAGENDR == 1, 'Male', 'Female')

df %>%
  group_by(gender) %>%
  summarize(water = mean(water)) %>%
  ggplot(aes(x = gender, y = water, fill = gender)) +
  geom_bar(stat = 'identity') +
  labs(x = 'Gender', y = 'Water Intake (ml)')
```

There doesn't seem to be a significant different between genders in water intake.

## Water Intake by Age

```{r}
ggplot(aes(x = RIDAGEYR, y = water), data = df) +
  geom_point() +
  labs(x = 'Age', y = 'Water Intake (ml)')
```

It doesn't look like age is a factor in water intake

## Race

```{r}
df <- df %>%
  mutate(
    race = case_when(
      RIDRETH3 == 1 | RIDRETH3 == 2 ~ 'Hispanic',
      RIDRETH3 == 3 ~ 'White',
      RIDRETH3 == 4 ~ 'Black',
      RIDRETH3 == 6 ~ 'Asian',
      RIDRETH3 == 7 ~ 'Other (or Multi-racial)'
    )
  )

df %>%
  group_by(race) %>%
  summarize(water = mean(water)) %>%
  ggplot(aes(x = race, y = water, fill = race)) +
    geom_bar(stat = 'identity') +
    labs(x = 'Race', y = 'Total Intake (ml)')
```

Water intake seems pretty consistent across races

## Education

```{r}
df2 <- df %>%
  filter(!is.na(DMDEDUC2))

df2 <- df2 %>%
  mutate(
    education = case_when(
      DMDEDUC2 == 1 ~ 'Less Than 9th Grade',
      DMDEDUC2 == 2 ~ '9th-12th Grade',
      DMDEDUC2 == 3 ~ 'High School Graduate/GED',
      DMDEDUC2 == 4 ~ 'Some College or AA Degree',
      DMDEDUC2 == 5 ~ 'College Graduate or Above'
    )
  )

df2 %>%
  group_by(education) %>%
  summarize(water = mean(water)) %>%
  ggplot(aes(x = education, y = water, fill = education)) +
    geom_bar(stat = 'identity') +
    scale_x_discrete(guide = guide_axis(n.dodge=3)) +
    labs(x = 'Education Level', y = 'Water Intake (ml)')
```

There doesn't seem to be a link between education level and water intake.

# Water Intake and BMI

There is an [established link](https://pmc.ncbi.nlm.nih.gov/articles/PMC4940461/) between water intake and BMI.

Let's see if this is the case with the most recent data.

## Linear Relationship

```{r}
df2 <- df %>%
  filter(!is.na(BMXBMI)) %>%
  filter(RIDAGEYR >= 18 & RIDAGEYR < 65) # Study was aged 18-64

ggplot(aes(x = BMXBMI, y = water), data = df2) +
  geom_point() +
  labs(x = 'BMI', y = 'Water Intake (ml)')
```

It's really hard to tell if there's a relationship, so let's try modeling it.

```{r}
model <- lm(BMXBMI ~ water, data = df2)
summary(model)
```

The model is not significant and the $R^2$ value is incredibly small.

## And Obesity

Another hypothesis might be whether low water intake is an indicator of obesity.

```{r}
df2$isObese <- ifelse(df2$BMXBMI > 30, 'Yes', 'No')

df2 %>%
  group_by(isObese) %>%
  summarize(water = mean(water)) %>%
  ggplot(aes(x = isObese, y = water, fill = isObese)) +
    geom_bar(stat = 'identity') +
    labs(x = 'Obese?', y = 'Water Intake (ml)')
```

This follows the lack of pattern seen above.

```{r}
df2$isObeseCode <- ifelse(df2$isObese == 'Yes', 1, 0)

model <- glm(isObeseCode ~ water, data = df2, family = 'binomial')
summary(model)
```

## Add Confounds

The model used in the paper cited at the beginning of the section used the confounds of `age`, `race`, `sex`, and `income-to-poverty ratio`. Let's see what the models look like when these confounds are added.

```{r}
library(effectsize)
library(car)

model <- lm(BMXBMI ~ water + RIDAGEYR + race + gender + INDFMMPC, data = df2)
summary(model)

anova_table <- Anova(model, type = 2)  # or type = 3 for Type III SS

# Calculate partial eta squared
eta_squared(anova_table, partial = TRUE)
```

Water intake is now a statistically significant predictor of BMI, but the effect size is very small.

This is particularly import to take into consideration because of how p-values are calculated: using the standard error - which is in turn dependent on n.

  * n here is 3171
  * n for the paper is 9528
  
# Mental Health

## Depression

For this we'll label depressed as having felt down, depressed, or hopeless several times or more over the last two weeks.

```{r}
df <- df %>%
  mutate(
    depression = case_when(
      DPQ020 == 0 ~ 'No',
      DPQ020 %in% c(1, 2, 3) ~ 'Yes',
      .default = 'remove'
    )
  )
df2 <- df %>%
  filter(depression != 'remove')

df2 %>%
  group_by(depression) %>%
  summarize(water = mean(water)) %>%
  ggplot(aes(x = depression, y = water, fill = depression)) +
    geom_bar(stat = 'identity') +
    labs(x = 'Depressed?', y = 'Average Water Intake (ml)')
```

On average, water intake does not seem to differ much between depressed and non-depressed individuals.

## Tired or Little Energy

For this we'll label tired as having been bothered by the following problems:  feeling tired or having little energy, split into 3 groups: no, somewhat frequently, and frequently.

```{r}
df <- df %>%
  mutate(
    tired = case_when(
      DPQ040 == 0 ~ 'No',
      DPQ040 == 1 ~ 'Somewhat Frequently',
      DPQ040 %in% c(2, 3) ~ 'Frequently',
      .default = 'remove'
    )
  )
df2 <- df %>%
  filter(tired != 'remove')

df2 %>%
  group_by(tired) %>%
  summarize(water = mean(water)) %>%
  ggplot(aes(x = tired, y = water, fill = tired)) +
    geom_bar(stat = 'identity') +
    labs(x = 'Tired or Little Energy?', y = 'Average Water Intake (ml)')
```

There doesn't seem to be much of a difference between groups.
